Constraints on Cosmic-Ray Particle Populations, Magnetic Fields, and Dark Matter from VHE Gamma-Ray Observations of the Coma Cluster of Galaxies with VERITAS

### 2011-08-30 ms.tex
Moved Fermi results table to come before constraints
Fermi analysis redone with 99% CL, text accordingly added in the main part and conclusions

Galaxy clusters have for many years been proposed as sources of gamma rays. If shock acceleration in the ICM is an efficent process, a population of highly relativistic cosmic-ray (CR) hadrons, in the form of protons and heavy ions, in the ICM is to be expected. The main energy-loss mechanism for CR hadrons at high energies is pion production through the interaction of CRs with nuclei in the ICM. Pions are short lived and decay; the charged pions to electrons and positrons,\footnote{A charged pion decays to a muon or antimuon, which subsequently decay to an electron or positron.} and the neutral pion to gamma rays. Due to the low density of the ICM ($n_{\mathrm{ICM}}\sim 10^{-3}$ cm$^{-3}$) and the sheer size of a cluster, CR hadrons will be confined in the cluster on very long timescales \citep[][]{article:Volk_etal:1996, article:Berezinsky_etal:1997} allowing them to accumulate over time.

@@

The Coma cluster of galaxies (ACO 1656) is one of the most thouroughly studied clusters over all wavelengths \citep{article:Voges_etal:1999}. Located at a distance of about 100 Mpc \citep[$z=0.023$;][]{article:StrubleRood:1999}, it is one of the closest and also one of the most massive clusters \citep[$M \sim 10^{15}M_{\odot}$;][]{article:Smith:1983, article:Kubo_etal:2008}. It hosts both a giant radio halo \citep{article:Giovannini_etal:1993,article:Thierbach_etal:2003} and peripheral radio relic, which appears connected to the radio halo with a ``diffuse'' bridge \citep[see discussion in][]{article:BrownRudnick:2010}. It has been suggested that the relic is an infall shock \citep{article:Ensslin_etal:1998} and succesively demonstrated by cosmological simulations that model the non-thermal emission processes \citep{article:PfrommerEnsslinSpringel:2008, article:Pfrommer:2008, article:Battaglia_etal:2009, article:Skillman_etal:2011}. Extended soft X-ray (SXR) emission is evident from the ROSAT all-sky survey in the 0.1 to 2.4 keV band \citep{article:BrielHenryBohringer:1992} and observations with XMM-Newton \citep{article:Briel_etal:2001} revealed substructure in the X-ray halo. As such it is a natural candidate for gamma-ray observations.

@@

Gamma-ray like events are separated from the CR background by imposing cuts on scaled parameters \citep{article:Aharonian_etal:1997, article:Krawczynski_etal:2006} calculated from a parameterized moment analysis of the shower images \citep{inproc:Hillas:1985}. The gamma-hadron separation cuts used in this analysis were optimized a priori for a weak point source (3\% Crab Nebula flux level) and a photon index of -2.4 using data taken on the Crab Nebula during the same epoch and are termed \emph{standard} cuts. Because the VHE gamma-ray spectrum for the Coma cluster is expected to be hard, photon index of about -2.3, these standard cuts are suitable for the analysis of the Coma cluster data set.

@@

There cosmological simulations only considered advective transport of CRs by turbulent gas motions which produces centrally enhanced profiles.  However, other means of CR transport such as diffusion and streaming may flatten the CR radial profiles. It has been suggested that advection is dominant only during mergers \citep{article:EnsslinPfrommerMiniatiSubramanian:2011}, in which case relaxed clusters could have a reduced gamma-ray luminosity by up to a factor of five compared with merging clusters. The goal of this and other future work of seeking for gamma-ray emission from clusters is to constrain the acceleration physics and transport properties of CRs.

We adopt the electron density profile for the Coma cluster that has been inferred from ROSAT X-ray observations \citep{article:BrielHenryBohringer:1992} and model the spatial and spectral distribution of CRs according to the model by \citet{article:PinzkePfrommer:2010} that neglects the contribution of cluster galaxies. Figure \ref{fig:spectrum} shows the expected spectrum of Coma, normalized to the total flux within the virial radius. [Add {\em Fermi} results]

@@

Table \ref{table:constraints} shows the expected gamma-ray fluxes within the regions of interest. For the most constraining regions tested (within an aperture of radius 0.2$^{\circ}$), those are a factor of 7.2 below the inferred upper limits (see Table \ref{table:results} and assuming a spectral index of $\alpha=2.3$ which matches the simulated one at energies $E=200$ GeV). Hence we scale the averaged relative CR pressure %$X_{\CR}$
% = \bra P_\CR \ket /\bra P_\rmn{th} \ket$ 
of our model by that factor, keep the spatial behavior and obtain an integrated limit of the relative CR pressure of $X_{\CR}<0.091$ within 0.2$^{\circ}$ that translates to a limit within the cluster virial radius of $X_\CR<0.13$ (solid lines of Figure \ref{fig:xcr}).

\subsection{Simplified analytical model}
To complement the simulation-inspired analysis and to facilitate comparison with earlier analytical work, we repeat the data analysis with a simplified analytical model that assumes a power-law CR spectrum and a constant relative CR pressure, i.e., we adopt the {\em isobaric model of CRs}. Since a priori, the CR spectral index is unconstrained, we vary it within plausible ranges of $2.1<\alpha_{\rmn{p}}<2.5$. This spectral range is compatible with the radio spectral index of the giant radio halo of Coma after accounting for the spectral
steepening at frequencies $\nu\sim5~\rmn{GHz}$ due to the Sunyaev-Zel'dovich effect \citep{article:PfrommerEnsslin:2004b}\footnote{Assuming a magnetic field of 1 $\mu$G, the CR protons responsible for the GHz radio emitting electrons have an energy of $\sim100$ GeV and are $\sim$ 20 times less energetic than those CR protons responsible for 200-GeV gamma-ray emission.}. Table \ref{table:constraints} shows the resulting constraints on $X_\CR$ which range between 0.05 and 1, strongly
depending on the assumed spectral index.

[Add {\em Fermi} results, critical spectral index]

@@

\subsection{Minimum gamma-ray flux}
For clusters that host radio halos we are able to derive a minimum gamma-ray flux in the hadronic model of CR interactions. The idea is based on the fact that a stationary distribution of CR electrons loses all its energy to synchrotron radiation for strong magnetic fields ($B \gg B_{\rmn{CMB}} \simeq 3.2 \mu$G) so that the ratio of gamma-ray to synchrotron flux becomes independent of the spatial distribution of CRs and thermal gas \citep{article:Voelk:1989, article:Pohl:1994}, in particular with $\alpha_{\nu}\simeq 1$ as the observed synchrotron spectral index.  Hence we can derive a minimum gamma-ray flux in the hadronic
model
\begin{equation}
\label{eq:Fmin}
\mathcal{F}_{\gamma,\rmn{min}} = \frac{\dps A_{\gamma}}{\dps A_{\nu}}\frac{\dps L_{\nu}}{\dps 4\pi D_{\rmn{lum}}^{2}},
\end{equation}
where $L_{\nu}$ is the observed luminosity of the radio mini-halo and $D_{\rmn{lum}}$ denotes the luminosity distance to the respective cluster. Lowering the magnetic field would require an increase in the energy density of CR electrons to reproduce the observed synchrotron luminosity and thus increase the associated gamma-ray flux.

To derive a minimum gamma-ray flux that is comparable to the upper limits, we need to determined the radio flux within the corresponding angular regions. To this end, we fit the point-source subtracted, azimuthally averaged radio halo profile at 1.38 GHz \citep{article:Deiss_etal:1997} with a $\beta$-model,
\begin{equation}
\label{beta}
 S_{\nu} (r_{\bot})= S_{0} \left[ 1 + \left( \frac{r_{\bot}}{r_{\rmn{c}}}\right)^{2}\right]^{-3\beta + 1/2},
\end{equation}
where $S_{0} = 1.1 \times 10^{-3}\,\rmn{Jy\,arcmin}^{-2}$, $r_{\rmn{c}} = 450$ kpc, and $\beta = 0.78$.  Within the error bars, this profile is consistent with 326-MHz data taken by \citet{article:Govoni_etal:2001} when scaled with a radio spectral index of 1.15.

The results for the minimum gamma-ray flux $F_{\rmn{min}}(>220~\rmn{GeV}$ and the minimum relative CR pressure, $X_{\CR,\,\rmn{min}} = F_{\rmn{min}}/F$ are shown in Table \ref{table:constraints}. Even in the most constraining case ($\alpha=2.1$, $<0.2^{\circ}$), these are a factor of about 60 below the upper limits obtained in this work.

[Add {\em Fermi} results]

\section{Constraining the Magnetic Field}
\label{sec:B}
We have seen in the previous section that we can obtain an absolute lower limit on the gamma-ray emission in the hadronic model by assuming high magnetic fields, $B>B_\rmn{CMB}$. Hence, we can turn this around by using our upper limit on the gamma-ray emission (and by extension on the CR pressure) to infer a lower limit on the magnetic field that is needed to explain the observed radio emission which, however, assumes the validity of the hadronic model. Lowering the gamma-ray constraint will tighten (increase) the magnetic field limit. Once this gets in conflict with magnetic field measurements by means of other methods, e.g., Faraday rotation measure (RM)\footnote{Generally, Faraday RM analysis of the magnetic-field strength by, e.g., background sources through clusters are degenerate with the magnetic coherence scale and may be biased by the unknown correlation between magnetic and density fluctuations.}, this would challenge the hadronic model of radio halos. The method that we use to constrain the magnetic field inherits a dependence on the assumed spatial structure that we parametrize by
\begin{equation}
\label{eq:B}
B(r) = B_{0} \,\left(\frac{n_{\rmn{e}(r)}}{n_{\rmn{e}(0)}}\right)^{\alpha_B},
\end{equation}
as suggested by Faraday RM studies and numerical magneto-hydrodynamical (MHD) simulations \citep[][and references therein]{article:Bonafede_etal:2010, article:Bonafede_etal:2011}. In fact, the magnetic field in the Coma cluster is among the best constrained, because its proximity permits RM observations of
seven radio-sources located at projected distances of 50 to 1500 kpc from the cluster center. The best-fit model obtains $B_{0} = 4.7^{+0.7}_{-0.8}\,\mu$G and $\alpha_{B} = 0.5^{+0.2}_{-0.1}$ \citep{article:Bonafede_etal:2010}. We aim for constraining the central field strength, $B_{0}$, which depends on the magnetic decline, $\alpha_{B}$, that we vary (within reasonable ranges of $\Delta\alpha_{B}=0.2$ as suggested from the Faraday RM studies) to address this dependence. We proceed as follows:

\begin{enumerate}
\item Given a model for the magnetic field, we determine the profile of the relative CR pressure, $X_{\CR}$, so that the hadronically produced CR electrons exactly matches the observed radio halo emission over the entire extent. To this end, we de-project the fit to the surface brightness profile (using an Abel integral equation, see Appendix of \citealt{article:PfrommerEnsslin:2004b}) yielding the radio emissivity,
\begin{equation}
\label{eq:Coma:radio}
j_{\nu} (r) = \frac{S_{0}}{2\pi\, r_{\rmn{c}}}\,
\frac{6\beta - 1}{\left(1 + r^{2}/r_{\rmn{c}}^{2}\right)^{3 \beta}}\,
\mathcal{B}\left(\frac{1}{2}, 3\beta\right)
= j_{\nu,0} \left(1 + r^2/r_{\rmn{c}}^{2}\right)^{-3 \beta},
\end{equation}
where $\mathcal{B}$ denotes the beta function. Using ``isobaric CR model'' to compute the secondary radio emission profile uniquely determines $X_\CR(r)$. It is generically true for weak magnetic fields ($B<B_{\rmn{CMB}}$) in the outer parts of the Coma halo that the product $X_{\CR}(r)X_{B}(r)$ has to increase by a factor of about 100 toward the radio halo periphery to account for the observed extent. If we were to adopt a steeper magnetic decline such as $\alpha_{B}=0.5$ which produces a flat $X_{B}(r)$, the relative CR pressure would have to rise accordingly by a factor of 100.

\item Given this realization for $X_{\CR}$, we compute the pion-decay gamma-ray surface brightness profile, integrate the flux within a radius of $\{0.13, 0.2, 0.4\}$ degree, and scale the CR profile in order to match the corresponding VERITAS upper limits. This scaling factor, $X_{\CR,0}$, depends on the CR spectral index, $\alpha$, (assuming a power-law CR population for simplicity) and the radial decline of the magnetic field, $\alpha_{B}$.

\item In the last step, we scale the secondary radio emission profile linearly with $X_{\CR,0}$ and solve for $B_{0}$ subject to the boundary condition to match the observed synchrotron profile. Note that for $B_{0} \gg B_{\rmn{CMB}}$ and radio spectral index, $\alpha_{\nu}$, the solution would be degenerate.
% according to Eqn. (\ref{eq:Lnu}). 
\end{enumerate}

Table~\ref{table:Bmin} shows the results on the lower limit of the central magnetic field that range from $B_{0} = 0.4$ to $1.2\,\mu$G. There are a few interesting remarks in order. (1) The hardest CR spectral indices are most constraining because of the tightest upper limits. (2) For a steeper magnetic decline (larger $\alpha_{B}$), the CR number density needs to be larger to match the observed radio emission profiles which would yield a higher gamma-ray flux so that the upper limits are more constraining. This implies tighter lower limits for $B_{0}$. However, the largest choice of $\alpha_B=0.7$, which represents a flux-freezing magnetic field, is possibly too steep in the outer parts as it neglects the contributions from turbulent dynamos in these dynamically active regions. (3) Interestingly, in all cases, we use the 0.4$^{\circ}$ limit that is always most constraining for our choice of magnetic field configurations.  This is because we need a substantially increasing relative CR pressure profile to match the observed radio profiles (for a given magnetic realization); hence that CR realization produces a larger flux within 0.4$^{\circ}$ in comparison to the ``isobaric CR model'' assumption ($X_{\CR} = \rmn{const.}$) for which the 0.2$^{\circ}$ upper limits are more constraining; hence the constraints on $B_{0}$ get better for the larger integration area. Physically, this may be achieved by considering CR transport processes such as streaming which produces an increasing $X_{\CR}$-profile due to the larger available phase space in the outer cluster regions.

As a final word, we note that the $X_{\CR}$ profiles always obey the energy condition, i.e., $P_{\CR} < P_{\mathrm{th}}$ over the entire range of the radio halo emission ($< 1$ Mpc) when assuming a central magnetic field of $B_{0}=3\,\mu$G that is at the lower 2-$\sigma$ interval bound inferred from Faraday RM
\citep{article:Bonafede_etal:2010}. Typical values at 1 Mpc range from $X_{\CR} = 0.02$
($\alpha_{B}=0.3$, $\alpha=2.1$) to $X_{\CR} = 1$ ($\alpha_{B}=0.7$,
$\alpha=2.5$).\footnote{See Figure 3 in \citet{article:PfrommerEnsslin:2004a} for the entire parameter range assuming minimum energy conditions and \citet{article:PfrommerEnsslin:2004b}, Figure 7 for a parametrization as adopted in this study.} We caution, however, that the minimum-energy condition is violated at the outer radio halo boundary for the range of minimum magnetic-field values inferred by this study.

@@

\section{Discussion}
VERITAS observed the Coma cluster of galaxies for a total of 18.6 hours of high-quality live time between March and May in 2008.

Various approaches to model CR population and magnetic field distribution are complementary (so far just ideas, need to be formulated):
\begin{itemize}
\item simplified ``isobaric CR model'': simple and widely used in the literature. A priori, physically not justified, old fashioned, should be replaced by the approached below.
\item simulation-based analytical approach: ``first principle approach'' that predicts the CR and B distribution for a given set of assumptions. Powerful since it only required the density profile as input, but limited in the modeled physics (no active CR transport) that appears to be necessary to explain the radio halo bimodality.
\item pragmatic approach as employed in \S \ref{sec:B}: model CR and B to reproduce observed radio emission profile. Minimum energy conditions (or equivalently equipartition arguments) represent a special case within this approach. A priori, also physically not justified, but quite powerful as it shows what the ``correct'' model has to achieve and can point into the direction what the relevant physics has to deliever. Compare CR transport vs. simulation inspired models
\end{itemize}

\acknowledgments
This research was supported by grants from the U.S. Department of Energy, the U.S. National Science Foundation, and the Smithsonian Institution, by NSERC in Canada, by Science Foundation Ireland, and by STFC in the UK. We acknowledge the excellent work of the technical support staff at the FLWO and the collaborating institutions in the construction and operation of the instrument.

>>>>

Galaxy clusters have for many years been proposed as sources of gamma rays. If shock acceleration in the ICM is an efficient process, a population of highly relativistic cosmic-ray (CR) hadrons, in the form of protons and heavy ions, in the ICM is to be expected. The main energy-loss mechanism for CR hadrons at high energies is pion production through the interaction of CRs with nuclei in the ICM. Pions are short lived and decay; the charged pions to electrons and positrons,\footnote{A charged pion decays to a muon or antimuon, which subsequently decay to an electron or positron.} and the neutral pion to gamma rays. Due to the low density of the ICM ($n_{\mathrm{ICM}}\sim 10^{-3}$ cm$^{-3}$) and the sheer size of a cluster, CR hadrons will be confined in the cluster on very long timescales \citep[][]{article:Volk_etal:1996, article:Berezinsky_etal:1997} allowing them to accumulate over time.

@@

The Coma cluster of galaxies (ACO 1656) is one of the most thoroughly studied clusters over all wavelengths \citep{article:Voges_etal:1999}. Located at a distance of about 100 Mpc \citep[$z=0.023$;][]{article:StrubleRood:1999}, it is one of the closest and also one of the most massive clusters \citep[$M \sim 10^{15}M_{\odot}$;][]{article:Smith:1983, article:Kubo_etal:2008}. It hosts both a giant radio halo \citep{article:Giovannini_etal:1993,article:Thierbach_etal:2003} and peripheral radio relic, which appears connected to the radio halo with a ``diffuse'' bridge \citep[see discussion in][]{article:BrownRudnick:2010}. It has been suggested that the relic is an infall shock \citep{article:Ensslin_etal:1998} and successively demonstrated by cosmological simulations that model the non-thermal emission processes \citep{article:PfrommerEnsslinSpringel:2008, article:Pfrommer:2008, article:Battaglia_etal:2009, article:Skillman_etal:2011}. Extended soft X-ray (SXR) emission is evident from the ROSAT all-sky survey in the 0.1 to 2.4 keV band \citep{article:BrielHenryBohringer:1992} and observations with XMM-Newton \citep{article:Briel_etal:2001} revealed substructure in the X-ray halo. As such it is a natural candidate for gamma-ray observations.

@@

Gamma-ray like events are separated from the CR background by imposing cuts on scaled parameters \citep{article:Aharonian_etal:1997, article:Krawczynski_etal:2006} calculated from a parametrized moment analysis of the shower images \citep{inproc:Hillas:1985}. The gamma-hadron separation cuts used in this analysis were optimized a priori for a weak point source (3\% Crab Nebula flux level) and a photon index of -2.4 using data taken on the Crab Nebula during the same epoch and are termed \emph{standard} cuts. Because the VHE gamma-ray spectrum for the Coma cluster is expected to be hard, photon index of about -2.3, these standard cuts are suitable for the analysis of the Coma cluster data set.

@@

These cosmological simulations only considered advective transport of CRs by turbulent gas motions which produces centrally enhanced profiles.  However, other means of CR transport such as diffusion and streaming may flatten the CR radial profiles. It has been suggested that advection is dominant only during mergers \citep{article:EnsslinPfrommerMiniatiSubramanian:2011}, in which case relaxed clusters could have a reduced gamma-ray luminosity by up to a factor of five compared with merging clusters. The goal of this and other future work of seeking for gamma-ray emission from clusters is to constrain the acceleration physics and transport properties of CRs.

We adopt the electron density profile for the Coma cluster that has been inferred from ROSAT X-ray observations \citep{article:BrielHenryBohringer:1992} and model the spatial and spectral distribution of CRs according to the model by \citet{article:PinzkePfrommer:2010} that neglects the contribution of cluster galaxies. Figure \ref{fig:spectrum} shows the expected spectrum of Coma, normalized to the total flux within the virial radius. In our further analysis, we use the gamma-ray flux limits of {\em Fermi} in the energy interval of 1-3 GeV which are most constraining.

@@

Table \ref{table:constraints} shows the expected gamma-ray fluxes within the regions of interest. For the most constraining regions tested (within an aperture of radius 0.2$^{\circ}$), those are a factor of 7.2 below the inferred upper limits of VERITAS (see Table \ref{table:results} and assuming a spectral index of $\alpha=2.3$ which matches the simulated one at energies $E=200$ GeV). Hence we scale the averaged relative CR pressure of our model by that factor, keep the spatial behavior and obtain an integrated limit of the relative CR pressure of $X_{\CR}<0.091$ within 0.2$^{\circ}$ that translates to a limit within the cluster virial radius of $X_\CR<0.13$ (solid lines of Figure \ref{fig:xcr}).

[Anders: add and discuss Fermi limits on XCR]

\subsection{Simplified analytical model}
To complement the simulation-inspired analysis and to facilitate comparison with
earlier analytical work, we repeat the data analysis with a simplified
analytical model that assumes a power-law CR spectrum and a constant relative CR
pressure, i.e., we adopt the {\em isobaric model of CRs}. Since a priori, the CR
spectral index is unconstrained, we vary it within plausible ranges of
$2.1<\alpha_{\rmn{p}}<2.5$. This spectral range is compatible with the radio
spectral index of the giant radio halo of Coma after accounting for the spectral
steepening at frequencies $\nu\sim5~\rmn{GHz}$ due to the Sunyaev-Zel'dovich
effect \citep{article:PfrommerEnsslin:2004b}\footnote{Assuming a magnetic field
  of 1 $\mu$G, the CR protons responsible for the GHz radio emitting electrons
  have an energy of $\sim100$ GeV and are $\sim$ 20 times less energetic than
  those CR protons responsible for 200-GeV gamma-ray emission.}. Table
\ref{table:constraints} shows the resulting constraints on $X_\CR$. The VERITAS
flux upper limits strongly depend on $\alpha$ due to the comparably large lever
arm from GeV energies (that dominate the CR pressure) to TeV-energies. The
limits within 0.2$^{\circ}$ are most constraining yielding limits on $X_\CR$
between 0.048 and 0.43 (for $\alpha$ varying between 2.1 and 2.5). The {\em
  Fermi} limits depend only weakly on $\alpha$ and are most constraining for an
aperture of 0.4$^{\circ}$.  The best limit on $X_\CR$ of 0.017 is achieved for
$\alpha=2.3$.

@@

\subsection{Minimum gamma-ray flux}
For clusters that host radio halos we are able to derive a minimum gamma-ray flux in the hadronic model of CR interactions. The idea is based on the fact that a stationary distribution of CR electrons loses all its energy to synchrotron radiation for strong magnetic fields ($B \gg B_{\rmn{CMB}} \simeq 3.2 \mu$G) so that the ratio of gamma-ray to synchrotron flux becomes independent of the spatial distribution of CRs and thermal gas \citep{article:Voelk:1989, article:Pohl:1994, article:Pfrommer:2008}, in particular with $\alpha_{\nu}\simeq 1$ as the observed synchrotron spectral index.  Hence we can derive a minimum gamma-ray flux in the hadronic
model
\begin{equation}
\label{eq:Fmin}
\mathcal{F}_{\gamma,\rmn{min}} = \frac{\dps A_{\gamma}}{\dps A_{\nu}}\frac{\dps L_{\nu}}{\dps 4\pi D_{\rmn{lum}}^{2}},
\end{equation}
where $L_{\nu}$ is the observed luminosity of the radio mini-halo,
$D_{\rmn{lum}}$ denotes the luminosity distance to the respective cluster,
$A_\gamma$ and $A_\nu$ are dimensional constants that depend on the hadronic
physics of the interaction \citep{2008MNRAS.385.1211P,
  2008MNRAS.385.1242P}. Lowering the magnetic field would require an increase in
the energy density of CR electrons to reproduce the observed synchrotron
luminosity and thus increase the associated gamma-ray flux.

To derive a minimum gamma-ray flux that is comparable to the upper limits, we need to determine the radio flux within the corresponding angular regions. To this end, we fit the point-source subtracted, azimuthally averaged radio halo profile at 1.38 GHz \citep{article:Deiss_etal:1997} with a $\beta$-model,
\begin{equation}
\label{beta}
 S_{\nu} (r_{\bot})= S_{0} \left[ 1 + \left( \frac{r_{\bot}}{r_{\rmn{c}}}\right)^{2}\right]^{-3\beta + 1/2},
\end{equation}
where $S_{0} = 1.1 \times 10^{-3}\,\rmn{Jy\,arcmin}^{-2}$, $r_{\rmn{c}} = 450$ kpc, and $\beta = 0.78$.  Within the error bars, this profile is consistent with 326-MHz data taken by \citet{article:Govoni_etal:2001} when scaled with a radio spectral index of 1.15.

The results for the minimum gamma-ray flux $F_{\rmn{min}}(>220~\rmn{GeV})$ and the minimum relative CR pressure, $X_{\CR,\,\rmn{min}} = F_{\rmn{min}}/F$ are shown in Table \ref{table:constraints}. Even in the most constraining cases, these are a factor of about 60 below the VERITAS upper limits (adopting $\alpha=2.1$, $<0.2^{\circ}$) and a factor of about 20 below the {\em Fermi} upper limits (adopting $\alpha=2.3$, $<0.4^{\circ}$).

\section{Constraining the Magnetic Field}
\label{sec:B}
We have seen in the previous section that we can obtain an absolute lower limit on the gamma-ray emission in the hadronic model by assuming high magnetic fields, $B>B_\rmn{CMB}$. Hence, we can turn this around by using our upper limit on the gamma-ray emission (and by extension on the CR pressure) to infer a lower limit on the magnetic field that is needed to explain the observed radio emission which, however, assuming the validity of the hadronic model \cite{article:PfrommerEnsslin:2004a}. Lowering the gamma-ray constraint will tighten (increase) the magnetic field limit. Once this gets in conflict with magnetic field measurements by means of other methods, e.g., Faraday rotation measure (RM)\footnote{Generally, Faraday RM analysis of the magnetic-field strength by, e.g., background sources through clusters are degenerate with the magnetic coherence scale and may be biased by the unknown correlation between magnetic and density fluctuations.}, this would challenge the hadronic model of radio halos. The method that we use to constrain the magnetic field inherits a dependence on the assumed spatial structure that we parametrize by
\begin{equation}
\label{eq:B}
B(r) = B_{0} \,\left(\frac{n_{\rmn{e}(r)}}{n_{\rmn{e}(0)}}\right)^{\alpha_B},
\end{equation}
as suggested by Faraday RM studies and numerical magneto-hydrodynamical (MHD) simulations \citep[][and references therein]{article:Bonafede_etal:2010, article:Bonafede_etal:2011}. In fact, the magnetic field in the Coma cluster is among the best constrained, because its proximity permits RM observations of
seven radio-sources located at projected distances of 50 to 1500 kpc from the cluster center. The best-fit model obtains $B_{0} = 4.7^{+0.7}_{-0.8}\,\mu$G and $\alpha_{B} = 0.5^{+0.2}_{-0.1}$ \citep{article:Bonafede_etal:2010}. We aim for constraining the central field strength, $B_{0}$, which depends on the magnetic decline, $\alpha_{B}$, that we vary (within reasonable ranges of $\Delta\alpha_{B}=0.2$ as suggested from the Faraday RM studies) to address this dependence. We proceed as follows:

\begin{enumerate}
\item Given a model for the magnetic field, we determine the profile of the relative CR pressure, $X_{\CR}(r)$, so that the hadronically produced CR electrons exactly matches the observed radio halo emission over the entire extent. To this end, we de-project the fit to the surface brightness profile (using an Abel integral equation, see Appendix of \citealt{article:PfrommerEnsslin:2004b}) yielding the radio emissivity,
\begin{equation}
\label{eq:Coma:radio}
j_{\nu} (r) = \frac{S_{0}}{2\pi\, r_{\rmn{c}}}\,
\frac{6\beta - 1}{\left(1 + r^{2}/r_{\rmn{c}}^{2}\right)^{3 \beta}}\,
\mathcal{B}\left(\frac{1}{2}, 3\beta\right)
= j_{\nu,0} \left(1 + r^2/r_{\rmn{c}}^{2}\right)^{-3 \beta},
\end{equation}
where $\mathcal{B}$ denotes the beta function. It is generically true for weak magnetic fields ($B<B_{\rmn{CMB}}$) in the outer parts of the Coma halo that the product $X_{\CR}(r)X_{B}(r)$ has to increase by a factor of about 100 toward the radio halo periphery to account for the observed extent. If we were to adopt a steeper magnetic decline such as $\alpha_{B}=0.5$ which produces a flat $X_{B}(r)$, the relative CR pressure would have to rise accordingly by a factor of 100.

\item Given this realization for $X_{\CR}$, we compute the pion-decay gamma-ray surface brightness profile, integrate the flux within a radius of $(0.13, 0.2, 0.4)$ degree, and scale the CR profile in order to match the corresponding VERITAS/{\em Fermi} upper limits. This scaling factor, $X_{\CR,0}$, depends on the CR spectral index, $\alpha$, (assuming a power-law CR population for simplicity) and the radial decline of the magnetic field, $\alpha_{B}$.

\item In the last step, we scale the secondary radio emission profile linearly with $X_{\CR,0}$ and solve for $B_{0}$ subject to the boundary condition to match the observed synchrotron profile. Note that for $B_{0} \gg B_{\rmn{CMB}}$ and a radio spectral index of $\alpha_{\nu}=1$, the solution would be degenerate.
\end{enumerate}

Table~\ref{table:Bmin} shows the results on the lower limit of the central
magnetic field that range for the VERITAS limits from $B_{0} = 0.4$ to
$1.2\,\mu$G and for the {\em Fermi} limits from $B_{0} = 1$ to $7\,\mu$G. Since
these lower limits on $B_0$ are for most of the parameter space spanned by
$\alpha_B$ and $\alpha$ below the values favored by Faraday RM, the hadronic
model is still viable. The only exception is the parameter combination of
$alpha_B=0.7$ and $\alpha=2.5$ which starts to be in (weak) tension with Faraday
RM data. This indicates the power of future gamma-ray observations of Coma in
constraining parameters of the hadronic model.

In addition there are a few interesting remarks in order. (1) For the VERITAS
limits, the hardest CR spectral indices are most constraining because of the
tightest upper limits. The opposite is true for the {\em Fermi} upper limits
which probe the region around the pion decay bump characterized by a shape
parameter that depends inversely on the CR spectral index
\cite{article:PfrommerEnsslin:2004b}.  (2) For a steeper magnetic decline
(larger $\alpha_{B}$), the CR number density needs to be larger to match the
observed radio emission profiles which would yield a higher gamma-ray flux so
that the upper limits are more constraining. This implies tighter lower limits
for $B_{0}$. However, the largest choice of $\alpha_B=0.7$, which represents a
flux-freezing magnetic field, is possibly too steep in the outer parts as it
neglects the contributions from turbulent dynamos in these dynamically active
regions with vorticity injection at corrugated structure formation shocks
\cite{article:PfrommerJones:2011}. (3) Interestingly, in all cases, we use the
0.4$^{\circ}$ limit that is always most constraining for our choice of magnetic
field configurations.  This is because we need a substantially increasing
relative CR pressure profile to match the observed radio profiles (for a given
magnetic realization); hence that CR realization produces a larger flux within
0.4$^{\circ}$ in comparison to the ``isobaric CR model'' assumption ($X_{\CR} =
\rmn{const.}$) for which the 0.2$^{\circ}$ upper limits are more constraining;
hence the constraints on $B_{0}$ get better for the larger integration
area. Physically, this may be achieved by considering CR transport processes
such as streaming which produces an increasing $X_{\CR}$-profile due to the
larger available phase space in the outer cluster regions.

As a final word, we note that the $X_{\CR}$ profiles always obey the energy
condition, i.e., $P_{\CR} < P_{\mathrm{th}}$ over the entire range of the radio
halo emission ($< 1$ Mpc) when assuming a central magnetic field of
$B_{0}=3\,\mu$G that is at the lower 2-$\sigma$ interval bound inferred from
Faraday RM \citep{article:Bonafede_etal:2010}.\footnote{Note that a central
  magnetic field of $3\,\mu$G corresponds in the Coma cluster to a relative
  magnetic energy density of $\tilde{X}_B=0.005$.} Typical values at 1 Mpc range
from $X_{\CR} = 0.02$ ($\alpha_{B}=0.3$, $\alpha=2.1$) to $X_{\CR} = 1$
($\alpha_{B}=0.7$, $\alpha=2.5$).\footnote{See Figure 3 in
  \citet{article:PfrommerEnsslin:2004a} for the entire parameter range assuming
  minimum energy conditions and \citet{article:PfrommerEnsslin:2004b}, Figure 7
  for a parametrization as adopted in this study.} We caution, however, that the
minimum-energy condition is violated at the outer radio halo boundary for the
range of minimum magnetic-field values inferred by this study.

@@

\section{Discussion and Conclusions}
VERITAS observed the Coma cluster of galaxies for a total of 18.6 hours of high-quality live time between March and May in 2008.

We employed various approaches to constrain the CR population and magnetic field
distribution that are complementary in their assumptions and hence well suited
in assessing underlying Bayesian priors in the models. (1) We use a simplified
``isobaric CR model'' that is characterized by a constant CR-to-thermal pressure
fraction and has a power-law momentum spectrum.  While this model is physically
not justified a priory, it is simple and widely used in the literature and
captures some aspects of more elaborate models such as (2) the simulation-based
analytical approach of \citet{article:PinzkePfrommer:2010}. This is a ``first
principle approach'' that predicts the CR distribution for a given set of
assumptions (namely the physics that was simulated). It is powerful since it
only requires the density profile as input due to the approximate universality
of the CR distribution (when neglecting CR diffusion and streaming which
appears, however, necessary to explain the radio halo bimodality). (3) We
finally use a pragmatic approach which models the CR and magnetic distribution
such as to reproduce the observed emission profile of the Coma radio halo. While
this approach is also physically not justified, it is powerful as it shows what
the ``correct'' model has to achieve and can point into the direction of the
relevant physics.

Within this pragmatic approach we employ two different directions. First,
adopting a high magnetic field everywhere in the cluster ($B\gg B_\rmn{CMB}$)
yields the minimum gamma-ray flux in the hadronic model of radio halos which we
find to be a factor of 20 (60) below the most constraining flux upper limits of
{\em Fermi} (VERITAS). Second, by matching the radio emission profile
(i.e. fixing the radial CR profile for a given magnetic field model) and by
requiring the pion-decay gamma-ray flux to match the flux upper limits
(i.e. fixing the normalization of the CR distribution) we obtain lower limits on
the magnetic field distribution under consideration. Our limits of the central
magnetic field range from $B_{0} = 0.4$ to $1.2\,\mu$G (for VERITAS limits) and
from $B_{0} = 1$ to $7\,\mu$G (for {\em Fermi} limits). Since these lower limits
on $B_0$ are below the values favored by Faraday RM, the hadronic model is still
viable.

Using the most constraining VERITAS limits, we can constrain the CR-to-thermal
pressure, $X_\CR$, to be below 0.048 and 0.43 (for a CR or gamma-ray spectral
index, $\alpha$, varying between 2.1 and 2.5).  The best {\em Fermi} flux limit
constrains $X_\CR$ to be below 0.017 (for $\alpha=2.3$), only weakly depending
on the assumed CR spectral index.

\acknowledgments
This research was supported by grants from the U.S. Department of Energy, the U.S. National Science Foundation, and the Smithsonian Institution, by NSERC in Canada, by Science Foundation Ireland, and by STFC in the UK. We acknowledge the excellent work of the technical support staff at the FLWO and the collaborating institutions in the construction and operation of the instrument. C.P. gratefully acknowledges financial support of the Klaus Tschira Foundation.
